---
title: "ALDEx2 CLR Benchmark"
author: "Benchmark Analysis"
date: "`r Sys.time()`"
format:
  html:
    toc: true
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---


```{r}
#| label: setup-R
#| include: false
#| echo: false
#| output: false

library(ALDEx2)
library(microbenchmark)
library(ggplot2)
library(grid)
library(bench)
library(ggbeeswarm)
library(testthat)
```

```{julia}
#| label: setup-Julia
#| include: false
#| echo: false
#| output: false

import Pkg

try
    using Plots
catch
    Pkg.add("Plots")
    using Plots
end

try
    using StatsPlots
catch
    Pkg.add("StatsPlots")
    using StatsPlots
end

try
    using BenchmarkTools
catch
    Pkg.add("BenchmarkTools")
    using BenchmarkTools
end
```


```{r}
#| label: load-data
#| echo: false
#| output: false

selex <- read.csv(
  "/Users/markpampuch/Dropbox/KAUST/PhD/ALDEx2_jl/ALDEx2_dev/TEST_DATA/selex_subset_1201-1600.csv",
  row.names = 1
)

conds <- c(rep("NS", 7), rep("S", 7))
```


```{r}
#| label: benchmark-R-native
#| cache: false
#| echo: false
#| output: false

set.seed(1)

bm_r <- mark(
  clr1 = aldex.clr(selex, conds, mc.samples = 128),
  iterations = 100
)

bm_r_ms <- as.numeric(bm_r$time[[1]])
```


```{r}
#| label: benchmark-Julia-to-R-interface
#| cache: false
#| echo: false
#| output: false

set.seed(1)

bm_julia_to_r <- mark(
  clr1 = aldex.clr(selex, conds, mc.samples = 128),
  iterations = 100
)

bm_julia_to_r_ms <- as.numeric(bm_julia_to_r$time[[1]])
```


```{julia}
#| label: benchmark-Julia
#| cache: false
#| echo: false
#| output: false

# Generate a random vector of size 1000
x = randn(1000)

# Define a mock function to do a basic operation (e.g., z-score normalization)
function mock_clr(x)
    μ = mean(x)
    σ = std(x)
    return (x .- μ) ./ σ
end

# Run a benchmark of the mock function
bm_julia = @benchmark mock_clr($x)

println(bm_julia)
display(median(bm_julia))

```

```{r}
#| label: plot-R-native
#| echo: false
#| output: false

# Move the ggplot2-based R and interface (bm_r_ms, bm_julia_to_r_ms) plotting code up here,
# but just *create* and assign the ggplot/gridExtra objects, don't display yet.

# Native ALDEx2 CLR, R version plots:
hist_plot <- ggplot(data.frame(x = bm_r_ms), aes(x = x)) +
  geom_histogram(bins = 100, fill = "steelblue", color = "black") +
  theme_minimal() +
  ggtitle("Native R ALDEx2 CLR Benchmark: Histogram")

# violin_plot_r_native <- autoplot(bm_r) + ggtitle("Violin Plot of Native ALDEx2 CLR, R version")

# Create a data frame from the numeric vector
df_plot_r_native <- data.frame(
  time_ms = bm_r_ms,
  expression = "bm_r"  # name of your benchmark
)

# Violin plot
# # violin_plot_r_native <- autoplot(bm_r) + ggtitle("Violin Plot of Native ALDEx2 CLR, R version")

violin_plot_r_native <- ggplot(df_plot_r_native, aes(x = expression, y = time_ms)) +
  geom_violin(fill = "skyblue", alpha = 0.5, color = "black") +
  geom_boxplot(width = 0.1, outlier.shape = NA) +          
  stat_summary(fun = median, geom = "point", color = "red", size = 2) +
  labs(
    title = "Violin Plot of Native R ALDEx2 CLR Benchmark",
    x = "Expression",
    y = "Execution Time (ms)" #, log10-scaled"
  ) +
  # scale_y_log10() +   # <-- log-scale the y-axis
  theme_minimal(base_size = 14)

r_native_grid <- gridExtra::arrangeGrob(
  violin_plot_r_native,
  hist_plot,
  ncol = 2
)
```

```{r}
#| label: plot-Julia-to-R-interface
#| echo: false
#| output: false

# Julia-to-R interface version plots:
hist_plot_julia_to_r <- ggplot(data.frame(x = bm_julia_to_r_ms), aes(x = x)) +
  geom_histogram(bins = 100, fill = "steelblue", color = "black") +
  theme_minimal() +
  ggtitle("ALDEx2 CLR, Julia to R interface version: Histogram")

# violin_plot_julia_to_r <- autoplot(bm_julia_to_r) + ggtitle("Violin Plot of ALDEx2 CLR, Julia to R interface version")

# Create a data frame from the numeric vector
df_plot_julia_to_r <- data.frame(
  time_ms = bm_julia_to_r_ms,
  expression = "bm_julia_to_r"  # name of your benchmark
)


violin_plot_julia_to_r <- ggplot(df_plot_julia_to_r, aes(x = expression, y = time_ms)) +
  geom_violin(fill = "skyblue", alpha = 0.5, color = "black") +
  geom_boxplot(width = 0.1, outlier.shape = NA) +          
  stat_summary(fun = median, geom = "point", color = "red", size = 2) +
  labs(
    title = "Violin Plot of ALDEx2 CLR, Julia to R interface version",
    x = "Expression",
    y = "Execution Time (ms)" #, log10-scaled"
  ) +
  # scale_y_log10() +   # <-- log-scale the y-axis
  theme_minimal(base_size = 14)

julia_to_r_grid <- gridExtra::arrangeGrob(
  violin_plot_julia_to_r,
  hist_plot_julia_to_r,
  ncol = 2
)

```

```{julia}
#| label: plot-julia
#| echo: false
#| output: false

# Extract timing data from benchmark
times_ns = bm_julia.times
times_ms = times_ns ./ 1_000_000  # Convert nanoseconds to milliseconds



# Create a violin plot of execution times using StatsPlots
# Create a horizontal violin plot
# The original `violin` call didn't plot because for 1D data you need a group.
# A working simple violin plot using StatsPlots:

# Violin plot
p_violin = violin(
    ["Julia Execution Time"],
    [times_ms],
    perm = false,            # still works
    xlabel = "",
    ylabel = "Time (ms)", #, log10-scaled",
    title = "ALDEx2 CLR, Julia Benchmark: Violin Plot",
    legend = false,
    # scale_y_log10 = false,
    color = :steelblue,
    linecolor = :black,
)

# Histogram
p_hist = histogram(
    times_ms,
    bins = 30,
    xlabel = "Time (milliseconds)",
    ylabel = "Frequency",
    title = "ALDEx2 CLR, Julia Benchmark: Histogram",
    color = :orange,
    linecolor = :black,
    alpha = 0.7
)
```

### All benchmark summaries

#### Native ALDEx2 CLR, R version summary
```{r}
#| label: summary-R-native
#| echo: false

summary(bm_r)
```

#### ALDEx2 CLR, Julia version summary
```{julia}
#| label: summary-Julia
#| echo: false

bm_julia
```

#### ALDEx2 CLR, Julia to R interface version summary
```{r}
#| label: summary-Julia-to-R
#| echo: false

summary(bm_julia_to_r)
```

### Native ALDEx2 CLR, R version benchmarks
```{r}
#| label: display-plots-R-native

grid::grid.draw(r_native_grid)
```

### ALDEx2 CLR, Julia to R interface version benchmarks
```{r}
#| label: display-plots-Julia-to-R-interface

grid::grid.draw(julia_to_r_grid)
```

### ALDEx2 CLR, Julia version benchmarks
```{julia}
#| label: display-plots-Julia
plot(p_violin, p_hist, layout = (1, 2))
```

### Test Custom Package Loading Into Benchmark Suite
```{r}
#| label: test-ALDEx2jlr
#| echo: true

# Load devtools if not already loaded
if (!require(devtools, quietly = TRUE)) {
  install.packages("devtools")
  library(devtools)
}

# Load ALDEx2jlr package from local development directory
# Using absolute path for reliability
aldex2jlr_path <- "/Users/markpampuch/Dropbox/KAUST/PhD/ALDEx2_jl/ALDEx2_dev/ALDEx2jlr"
devtools::load_all(aldex2jlr_path)

# Setup ALDEx2.jl (this initializes Julia and loads the package)
aldex2jl <- aldex2jl_setup(pkg_check = TRUE)

# Call hello_world function
result <- aldex2jl$hello_world()
print(result)

# Call hello_world! function (note: ! is converted to _bang in R)
result_bang <- aldex2jl$hello_world_bang()
print(result_bang)
```


```{julia}
#| label: test-ALDEx2.jl
#| echo: true

# Load the development package
import Pkg

# Use absolute path to the package directory (contains Project.toml)
package_path = "/Users/markpampuch/Dropbox/KAUST/PhD/ALDEx2_jl/ALDEx2_dev/ALDEx2.jl/ALDEx2";

# Add the package in development mode
Pkg.develop(path = package_path)

# Import the package
using ALDEx2

# Call the hello_world function
hello_world()
```

### Check Source Codes
```{julia}
#| label: test-ALDEx2.jl-source-code
#| echo: true

# Read and display the source code of ALDEx2.jl
source_file = "/Users/markpampuch/Dropbox/KAUST/PhD/ALDEx2_jl/ALDEx2_dev/ALDEx2.jl/ALDEx2/src/ALDEx2.jl";
println("Source code of ALDEx2.jl:");
println("=" ^ 60);
read(source_file, String) |> print
```

```{r}
#| label: test-ALDEx2.jl-source-code-R
#| echo: true

# Read and display the source code of ALDEx2.jl
print(aldex2jl$hello_world)
print(aldex2jl$hello_world_bang)
```

### Run Unit Tests
```{r}
#| label: run-unit-tests-Julia-to-R-interface
#| echo: true

mywd <- getwd()

# Run unit tests of ALDEx2.jl
setwd("/Users/markpampuch/Dropbox/KAUST/PhD/ALDEx2_jl/ALDEx2_dev/ALDEx2jlr")
devtools::test()
setwd(mywd)

```

### Output the git branch and commit hash
```{r}
#| label: git-branch-and-commit-hash
#| echo: true

git_branch <- system("git branch --show-current", intern = TRUE)
git_commit <- system("git rev-parse HEAD", intern = TRUE)
print(git_branch)
print(git_commit)
```